---
title: "Lab9"
author: "J Bai"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(microbenchmark)
library(profvis)
```

# Activity 9
## 1. Consider a function that counts number of odd numbers in a given vector
```{r}
odd_count <- function(x) { 
  odd_num = 0
  for (i in 1:length(x)) {
    if (x[i] %% 2 == 1) odd_num = odd_num + 1 
  }
  return(odd_num) 
}

odd_count_v <- function(x) {
  return(sum(x %% 2))
}

x <- c(1, 2, 4, 6, 9, 7, 11, 28)
odd_count(x)
odd_count_v(x)

microbenchmark(odd_count(x), odd_count_v(x))
```


## 2. Create a matrix and calculate sum of each column
```{r}
test.matrix <- matrix(1:9, nrow = 3)
test.matrix

# loop
#colSumsLoop <- function(m) {
#  result <- rep(0, ncol(m))
#  for(i in 1:ncol(m)) {
#    result[i] <- sum(m[, i])
#  }
#  return(result)
#}

colSumsLoop <- function(A) {
  m <- nrow(A)
  n <- ncol(A)
  y <- rep(0, n)
  for(j in 1:n) {
    for(i in 1:m) {
      y[j] = y[j] + A[i,j]
    }
  }
  return(y)
}
colSumsLoop(test.matrix)

# apply
colSumsApply <- function(m) {
  apply(m, 2, sum)
}

colSumsApply(test.matrix)

# colSums
colSums(test.matrix)

microbenchmark(colSumsLoop(test.matrix), colSumsApply(test.matrix), colSums(test.matrix))

```

## 3. Create a random vector of positive integers of length 1000000 elements. Use profvis() function to profile the odd_count() function defined in the problem 1 with the vector as an input.
```{r}
vector.million <- sample.int(100, 1000000, replace = TRUE)
profvis({
  odd_count <- function(x) { 
    odd_num = 0
    for (i in 1:length(x)) {
      if (x[i] %% 2 == 1) odd_num = odd_num + 1 
    }
    return(odd_num) 
  }
  odd_count(vector.million)

  odd_count_v <- function(x) {
    return(sum(x %% 2))
  }
  odd_count(vector.million)
})
```

# Assignment Lab 9
## 2 Modify the sorting function (sort_vec) from “Assignment 8” (problem 3) so that it should take an additional argument ascending which causes sorting in increasing order when ‘ascending = TRUE’. 
```{r}
sort_vec <- function(x, ascending=TRUE) {
  if (length(x) < 2) 
    return(x)
  for(last in length(x):2) { 
    for(first in 1:(last - 1)) {
      if(ascending) {
        if(x[first] > x[first + 1]) { 
          temp = x[first]
          x[first] = x[first + 1] 
          x[first + 1] = temp
        }
      }
      else {
        if(x[first] <= x[first + 1]) { 
          temp = x[first]
          x[first] = x[first + 1] 
          x[first + 1] = temp
        }
      }
    }
  }
  return(x) 
}
```

## 3. Consider a simple random walk with starting point 0 and a step -1 or 1. Below is the code with dynamically allocated memory. Write your code with preallocated memory and compare time for both versions using system.time() function (use N = 1000, 10000 and 1000000).
```{r}
N <- 100000
data_series = 0 
system.time(
  {for (i in 2:N) {
    data_series[i] = data_series[i-1] + sample(c(-1, 1), 1) }
  })


data_series <- rep(0,N)
system.time(
  {
    for (i in 2:N) {
    data_series[i] = data_series[i-1] + sample(c(-1, 1), 1) }
  })
```













